---
title: "readepi-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{readepi-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

You can install the development version of readepi from
[GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
devtools::install_github("Karim-Mane/readepi")
suppressPackageStartupMessages(library(readepi))
```

# Reading data from file or directory
## importing data from JSON file
```{r eval=FALSE}
file = system.file("extdata", "test.json", package = "readepi")
data = readepi(file.path = file)
```

## importing data from excel file with 2 sheets
here we are importing from the second excel sheet
```{r eval=FALSE}
file = system.file("extdata", "test.xlsx", package = "readepi")
data = readepi(file.path = file, which = "Sheet2")
```

## importing data from several files in a directory
```{r eval=FALSE}
# reading all files in the given directory
dir.path = "inst/extdata"
data = readepi(file.path = dir.path)

# reading only txt files
data = readepi(file.path = dir.path, pattern = "txt")
```

# Reading data from relational database management systems (RDBMS): HDSS, EMRS, REDCap
## importing data from RDBMS into R
### importing from REDCap
The user must have read access to the system where the data will be read from. The user credential must be stored in a tab-delimited file that is only accessible to him/her. To see the structure of the credential file, use the `show_example_file()` function.    
we were given access to the **Pats__Covid_19_Cohort_1_Screening**.    
```{r eval=FALSE}
# credentials.file = "/Users/karimmane/Documents/Karim/LSHTM/Data/.credentials_karim.ini"
# display the structure of the credentials file
# show_example_file()

# reading all fields and records the project
data = readepi(credentials.file, project.id="Pats__Covid_19_Cohort_1_Screening")
project.data = data$data
project.metadeta = data$metadata

# reading specified fields and all records from the project
fields = c("day_1_q_ran_id","redcap_event_name","day_1_q_1a","day_1_q_1b","day_1_q_1c","day_1_q_1","day_1_q_2","day_1_q_3","day_1_q_4","day_1_q_5")
data = readepi(credentials.file, project.id="Pats__Covid_19_Cohort_1_Screening", fields = fields)

# reading specified records and all fields from the project
records = c("C10001/3","C10002/1","C10003/7","C10004/5","C10005/9","C10006/8","C10007/6","C10008/4","C10009/2","C10010/1")
data = readepi(credentials.file, project.id="Pats__Covid_19_Cohort_1_Screening", records =  records)

# reading specified records and fields from the project
data = readepi(credentials.file, project.id="Pats__Covid_19_Cohort_1_Screening", records =  records, fields = fields)
```

### importing from HDSS or EMRS
we were given access to the **IBS_BHDSS** database.   
```{r eval=FALSE}
# reading all fields and all records from the `dss_events` table
data = readepi(credentials.file, project.id="IBS_BHDSS", driver.name = "ODBC Driver 17 for SQL Server", table.name = "dss_events")

# reading specified fields and all records from the table
data = readepi(credentials.file, project.id="IBS_BHDSS", driver.name = "ODBC Driver 17 for SQL Server", table.name = "dss_events", fields = fields)

# reading specified records and all fields from the table
data = readepi(credentials.file, project.id="IBS_BHDSS", driver.name = "ODBC Driver 17 for SQL Server", table.name = "dss_events", records = records)

# reading specified fields and records from the table
data = readepi(credentials.file, project.id="IBS_BHDSS", driver.name = "ODBC Driver 17 for SQL Server", table.name = "dss_events", records = records, fields = fields)
```

